Fractional Types

### 2016-12-18 typespp.tex
opsem
Fix references to opsem

\begin{figure}[t]
{\footnotesize
\[\begin{array}{cc}
\begin{array}[t]{rlcl}
\evalone{\unitepl}{&(\inr{v})} &=& v \\
\evalone{\unitipl}{&v} &=& \inr{v} \\
\evalone{\unitepr}{&(\inl{v})} &=& v \\
\evalone{\unitipr}{&v} &=& \inl{v} \\
\evalone{\swapp}{&(\inl{v})} &=& \inr{v} \\
\evalone{\swapp}{&(\inr{v})} &=& \inl{v} \\
\evalone{\assoclp}{&(\inl{v})} &=& \inl{(\inl{v})} \\
\evalone{\assoclp}{&(\inr{(\inl{v})})} &=& \inl{(\inr{v})} \\
\evalone{\assoclp}{&(\inr{(\inr{v})})} &=& \inr{v} \\
\evalone{\assocrp}{&(\inl{(\inl{v})})} &=& \inl{v} \\
\evalone{\assocrp}{&(\inl{(\inr{v})})} &=& \inr{(\inl{v})} \\
\evalone{\assocrp}{&(\inr{v})} &=& \inr{(\inr{v})}
\end{array} &
\begin{array}[t]{rlcl}
\evalone{\unitetl}{&(\unitv,v)} &=& v \\
\evalone{\unititl}{&v} &=& (\unitv,v) \\
\evalone{\unitetr}{&(v,\unitv)} &=& v \\
\evalone{\unititr}{&v} &=& (v,\unitv) \\
\evalone{\swapt}{&(v_1,v_2)} &=& (v_2,v_1) \\
\evalone{\assoclt}{&(v_1,(v_2,v_3))} &=& ((v_1,v_2),v_3) \\
\evalone{\assocrt}{&((v_1,v_2),v_3)} &=& (v_1,(v_2,v_3))
\end{array} \\
\\
\begin{array}[t]{rlcl}
\evalone{\absorbr}{&(v,\_)} &=& v \\
\evalone{\absorbl}{&(\_,v)} &=& v \\
\evalone{\dist}{&(\inl{v_1},v_3)} &=& \inl{(v_1,v_3)} \\
\evalone{\dist}{&(\inr{v_2},v_3)} &=& \inr{(v_2,v_3)} \\
\evalone{\factor}{&\inl{(v_1,v_3)}} &=& (\inl{v_1},v_3) \\
\evalone{\factor}{&\inr{(v_2,v_3)}} &=& (\inr{v_2},v_3) \\
\evalone{\distl}{&(v_1,\inl{v_3})} &=& \inl{(v_1,v_3)} \\
\evalone{\distl}{&(v_2,\inr{v_3})} &=& \inr{(v_2,v_3)} \\
\evalone{\factorl}{&\inl{(v_1,v_3)}} &=& (v_1,\inl{v_3}) \\
\evalone{\factorl}{&\inr{(v_2,v_3)}} &=& (v_2,\inr{v_3})
\end{array} &
\begin{array}[t]{rlcl}
\evalone{\idiso}{&v} &=& v \\
\evalone{(c_1 \odot c_2)}{&v} &=&
  \evalone{c_2}{(\evalone{c_1}{v})} \\
\evalone{(c_1 \oplus c_2)}{&(\inl{v})} &=&
  \inl{(\evalone{c_1}{v})} \\
\evalone{(c_1 \oplus c_2)}{&(\inr{v})} &=&
  \inr{(\evalone{c_2}{v})} \\
\evalone{(c_1 \otimes c_2)}{&(v_1,v_2)} &=&
  (\evalone{c_1}v_1, \evalone{c_2}v_2)
\end{array}
\end{array}\]
\caption{\label{opsem}$\Pi$ operational semantics}
}
\end{figure}

@@

We give an operational semantics for the 1-combinators of $\Pi$ which
represent the conventional layer of programs.  Operationally, the
semantics consists of a pair of evaluators that
take a combinator and a value and propagate the value in the forward
direction~$\triangleright$ or in the backward
direction~$\triangleleft$. We show the complete forward evaluator in
Fig.~\ref{opsem}; the backward evaluator is easy to infer.

@@

The operational semantics for all the primitive combinators is a
simple transliteration of Fig.~\ref{opsem}. We omit the
straightforward implementation of prim and its inverse
below:

@@

>>>>

\begin{figure}[t]
{\footnotesize
\[\begin{array}{cc}
\begin{array}[t]{rlcl}
\evalone{\unitepl}{&(\inr{v})} &=& v \\
\evalone{\unitipl}{&v} &=& \inr{v} \\
\evalone{\unitepr}{&(\inl{v})} &=& v \\
\evalone{\unitipr}{&v} &=& \inl{v} \\
\evalone{\swapp}{&(\inl{v})} &=& \inr{v} \\
\evalone{\swapp}{&(\inr{v})} &=& \inl{v} \\
\evalone{\assoclp}{&(\inl{v})} &=& \inl{(\inl{v})} \\
\evalone{\assoclp}{&(\inr{(\inl{v})})} &=& \inl{(\inr{v})} \\
\evalone{\assoclp}{&(\inr{(\inr{v})})} &=& \inr{v} \\
\evalone{\assocrp}{&(\inl{(\inl{v})})} &=& \inl{v} \\
\evalone{\assocrp}{&(\inl{(\inr{v})})} &=& \inr{(\inl{v})} \\
\evalone{\assocrp}{&(\inr{v})} &=& \inr{(\inr{v})}
\end{array} &
\begin{array}[t]{rlcl}
\evalone{\unitetl}{&(\unitv,v)} &=& v \\
\evalone{\unititl}{&v} &=& (\unitv,v) \\
\evalone{\unitetr}{&(v,\unitv)} &=& v \\
\evalone{\unititr}{&v} &=& (v,\unitv) \\
\evalone{\swapt}{&(v_1,v_2)} &=& (v_2,v_1) \\
\evalone{\assoclt}{&(v_1,(v_2,v_3))} &=& ((v_1,v_2),v_3) \\
\evalone{\assocrt}{&((v_1,v_2),v_3)} &=& (v_1,(v_2,v_3))
\end{array} \\
\\
\begin{array}[t]{rlcl}
\evalone{\absorbr}{&(v,\_)} &=& v \\
\evalone{\absorbl}{&(\_,v)} &=& v \\
\evalone{\dist}{&(\inl{v_1},v_3)} &=& \inl{(v_1,v_3)} \\
\evalone{\dist}{&(\inr{v_2},v_3)} &=& \inr{(v_2,v_3)} \\
\evalone{\factor}{&\inl{(v_1,v_3)}} &=& (\inl{v_1},v_3) \\
\evalone{\factor}{&\inr{(v_2,v_3)}} &=& (\inr{v_2},v_3) \\
\evalone{\distl}{&(v_1,\inl{v_3})} &=& \inl{(v_1,v_3)} \\
\evalone{\distl}{&(v_2,\inr{v_3})} &=& \inr{(v_2,v_3)} \\
\evalone{\factorl}{&\inl{(v_1,v_3)}} &=& (v_1,\inl{v_3}) \\
\evalone{\factorl}{&\inr{(v_2,v_3)}} &=& (v_2,\inr{v_3})
\end{array} &
\begin{array}[t]{rlcl}
\evalone{\idiso}{&v} &=& v \\
\evalone{(c_1 \odot c_2)}{&v} &=&
  \evalone{c_2}{(\evalone{c_1}{v})} \\
\evalone{(c_1 \oplus c_2)}{&(\inl{v})} &=&
  \inl{(\evalone{c_1}{v})} \\
\evalone{(c_1 \oplus c_2)}{&(\inr{v})} &=&
  \inr{(\evalone{c_2}{v})} \\
\evalone{(c_1 \otimes c_2)}{&(v_1,v_2)} &=&
  (\evalone{c_1}v_1, \evalone{c_2}v_2)
\end{array}
\end{array}\]
\caption{\label{fig:opsem}$\Pi$ operational semantics}
}
\end{figure}

@@

We give an operational semantics for the 1-combinators of $\Pi$ which
represent the conventional layer of programs.  Operationally, the
semantics consists of a pair of evaluators that
take a combinator and a value and propagate the value in the forward
direction~$\triangleright$ or in the backward
direction~$\triangleleft$. We show the complete forward evaluator in
Fig.~\ref{fig:opsem}; the backward evaluator is easy to infer.

@@

The operational semantics for all the primitive combinators is a
simple transliteration of Fig.~\ref{fig:opsem}. We omit the
straightforward implementation of prim and its inverse
below:

@@

\begin{figure}[t]
{\footnotesize
\[\begin{array}{cc}
\begin{array}[t]{rlcl}
  \evalone{\unitetl/}{&\langle id^{k},\mathbb{V} \rangle} &=& \mathbb{V} \\
  \evalone{\unititl/}{& \mathbb{V}} &=& \langle id^{k_1},\mathbb{V} \rangle \\
  \evalone{\unitetr/}{&\langle \mathbb{V},id^{k} \rangle} &=& \mathbb{V} \\
  \evalone{\unititr/}{& \mathbb{V}} &=& \langle \mathbb{V},id^{k} \rangle\\
  \evalone{\swapt}{&\langle \mathbb{V}_1, \mathbb{V}_2 \rangle} &=& \langle \mathbb{V}_2, \mathbb{V}_1 \rangle \\
  \evalone{\assoclt/}{&\langle c^k , \langle 1/c^{k}, c^{k} \rangle\rangle} &=& \langle \langle c^k , 1/c^{k} \rangle, c^{k} \rangle \\
  \evalone{\assocrt/}{&\langle \langle c^k , 1/c^{k} \rangle, c^{k} \rangle} &=& \langle c^k , \langle 1/c^{k}, c^{k} \rangle\rangle \\
\end{array}
&
\begin{array}[t]{rlcl}
  \evalone{\eta-}{& id^{k_1}} &=& \langle 1/c^{k_2},c^{k_2} \rangle\\
  \evalone{\eta+}{& id^{k_1}} &=& \langle c^{k_2},1/c^{k_2} \rangle\\
  \evalone{\epsilon-}{& \langle 1/c^{k_1},c^{k_1} \rangle} &=& id^{k_2} \\
  \evalone{\epsilon+}{& \langle c^{k_1},1/c^{k_1} \rangle} &=& id^{k_2} \\
\end{array}

\end{array}\]
\caption{\label{fig:opsem-frac}$\Pi^/$ operational semantics}
}
\end{figure}
